---
title: "R Notebook"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
```

```{r Creating The Old Boring League Table}
pl <- read.csv("results/1819.csv")
pl <- pl[c(1:8)]

a <- pl %>% group_by(HomeTeam) %>% count(FTR) %>% spread(key = FTR, value= n)
b <- pl %>% group_by(AwayTeam) %>% count(FTR) %>% spread(key = FTR, value= n)

colnames(a) <- c("Team","L","D","W")
colnames(b) <- c("Team","W","D","L")

a <- a[c(1,4,3,2)]
a <- a %>% replace(is.na(a),0)
b <- b %>% replace(is.na(b),0)

c <- bind_rows(a, 
          b) %>% 
  group_by(Team) %>% 
  summarise_at(c("W","D","L"),sum)

c <- c %>% mutate(pts=c(W*3+D*1+L*0))
c <- arrange(c,desc(pts))

d <- pl %>% group_by(HomeTeam) %>% summarise(GF=sum(FTHG), GA=sum(FTAG))
e <- pl %>% group_by(AwayTeam) %>% summarise(GF=sum(FTAG), GA=sum(FTHG))

colnames(d) <- c("Team","GF","GA")
colnames(e) <- c("Team","GF","GA")

f <- bind_rows(d, 
          e) %>% 
  group_by(Team) %>% 
  summarise_at(c("GF","GA"),sum)

table <- left_join(c,f,by="Team")
```

Must install Cairo through terminal in order to install gdtools in order to install hrbrthemes
- brew install cairo

```{r Now For The Good Stuff - Setup}
library(gganimate)
library(hrbrthemes)
library(lubridate)
library(zoo)
library(tweenr)

hrbrthemes::import_roboto_condensed()
```

```{r Now For The Good Stuff - By Week - Data}
rm(list=ls())
data <- read.csv("results/1819.csv")
data$Date <- dmy(data$Date)

data$HT_pts <- ifelse(data$FTR=="H",3,ifelse(data$FTR=="D",1,0))
data$AT_pts <- ifelse(data$FTR=="A",3,ifelse(data$FTR=="D",1,0))

a <- data %>% select(Date,HomeTeam,HT_pts)
b <- data %>% select(Date,AwayTeam,AT_pts)

colnames(a) <- c("date","team","pts")
colnames(b) <- c("date","team","pts")

c <- rbind(a,b)
c <- arrange(c,date)
c <- c %>% group_by(team) %>% mutate(cumpts=cumsum(pts))

c <- c %>% group_by(team) %>% mutate(week=rep(1:38))

c <- c %>% group_by(week) %>% arrange(desc(cumpts)) %>% mutate(rank=1:20)
```

```{r Now For The Good Stuff - By Week - The Good Stuff}
p <- c %>%
    ggplot(aes(x = -rank,y = cumpts)) +
        geom_tile(aes(y = cumpts / 2, height = cumpts, fill=team), width = 0.9) +
        geom_text(aes(label = team), hjust = "right", colour = "black", fontface = "bold") +
        geom_text(aes(label = scales::comma(cumpts)), hjust = "left", colour = "grey30") +
        coord_flip(clip="off") + 
        scale_x_discrete("") + 
        scale_y_continuous("",labels=scales::comma) +
        hrbrthemes::theme_ipsum(plot_title_size = 32, subtitle_size = 24, caption_size = 20, base_size = 20) +
        theme(panel.grid.major.y=element_blank(),
              panel.grid.minor.x=element_blank(),
              legend.position = c(0.4, 0.2),
              plot.margin = margin(1,1,1,2,"cm"),
              axis.text.y=element_blank()) +
# gganimate code to transition by year:
        transition_time(week) +
        ease_aes('linear') + 
        labs(title='Premier League Standings',
             subtitle='Points in Week {round(frame_time,0)}')

anim <- animate(p, width = 1200, height = 900)

anim_save("out5.gif", anim)
```

Performing the visualization by date is actually a much better representation of the feel of the title chase. As matchweeks get shifted due to cup obligations forcing rescheduling of matches, using date actually highlights that while gaps may a matter of points in matchweeks, they feel significantly more broad as teams get their matchweeks completed early.

```{r Now For The Good Stuff - By Day - Data}
rm(list=ls())
data <- read.csv("results/1819.csv")
data$Date <- dmy(data$Date)

data$HT_pts <- ifelse(data$FTR=="H",3,ifelse(data$FTR=="D",1,0))
data$AT_pts <- ifelse(data$FTR=="A",3,ifelse(data$FTR=="D",1,0))

a <- data %>% select(Date,HomeTeam,HT_pts)
b <- data %>% select(Date,AwayTeam,AT_pts)

colnames(a) <- c("date","team","pts")
colnames(b) <- c("date","team","pts")

c <- rbind(a,b)
c <- arrange(c,date)
c <- c %>% group_by(team) %>% mutate(cumpts=cumsum(pts))

date_list <- data.frame(date=seq(min(c$date),max(c$date),1))

date_team_comb <- merge(unique(c$team),date_list,all=T)

# Still gotta figure out how to do the dates properly. But each team needs each date represented with the total cumulative points upto that date. Once that is complete, it will be easy to created the updated rankings in Line 136 by switching group_by(week) to group_by(date).


c <- c %>% group_by(date) %>% arrange(desc(cumpts)) %>% mutate(rank=1:20)
```

